SPP-Perc: Peceptron Learning Based Enhanced Signature Path Prefetcher

### 2018-12-02 design_arch.tex
spelling
some comments

As discussed in Section~\ref{sec:Background-Perceptron}, perceptrons
are useful, light-weight mechanisms to pull together disparate forms
of information and synthesize a decision from them.  In this case, we
will leverage a variety of different relative characteristics of a
given prefetch, such as speculation depth, page address and offset,
into a binary useful/non-useful prediction.  In the remainder of this
section, we discus the the general structure of PPF for enhancing a
given underlying prefetcher.  We will show that PPF is a generalized
prefetch filtering mechanism which may adapted to any prefetcher with
minimal modifications. %Gino: Are these modifications minimal?

\subsection{The Perceptron Filter}
\label{Arch-Perceptron}

The Perceptron Filter is organized as tables of perceptron weights,
with a separate table of weights for each feature.  A different number
of bits are used for each feature to index into its corresponding
table.
%Gino: repharse below?
Each feature's value is used to index into the corresponding table.
This indexing is done using a different number of bits for each
feature.  At most, 12 bits of a feature are used to index into the
weights table. Certain features require more resolution \textit{i.e.},
the full 12 bits of indexing.  On the other hand, some features
requiring lower resolution require as few as 7 bits of indexing.  The
variable indexing was determined by studying the features and
fine-tuned empirically to achieve a good accuracy vs hardware overhead
trade-off.  Exact details are mentioned under ``Area Overheads'' in
Section~\ref{Method-Overheads}

A single entry in the table corresponds to a perceptron weight.  Each
weight is a 5 bit counter - saturating at -16 and +15 and initialized
to zero at the beginning of a program's execution. Our proposed design
uses nine features, represented in nine perceptron weight tables.

\textbf{Inference}\newline In the ChampSim simulator, the prefetching
method is invoked on every L2 Cache demand access.  At that point, the
base prefetcher has the option to perform a prefetch -- if it does,
then it has a choice of how many cache lines to prefetch.  The
suggested prefetches can be either placed in the L2 cache or the
last-level cache based on the prefetcher's internal confidence
mechanism.

@@

Once all the weights are retrieved, they are summed and compared to a
preset threshold ,PERC\_ THRESHOLD\_LO.  Only the prefetch candidates
with perceptron sum higher than the threshold qualify for prefetching.
The prefetches that qualify through the perceptron stage are recorded
in the ``Prefetch Table". The prefetch table is a 1024-entry, direct
mapped structure taht conatains all metadata required to recreate the
features.  %Gino: Recreate the features?
When the feedback of the current prefetched line is available at a
later stage, the stored data is used to train the perceptron.

The sum of the perceptron weights are also used to decide whether to
place teh data in the L2 or the last level cache. All prefetch
candidates that are qualified until this stage are compared against
PERC\_THRESHOLD\_HI to decide the fill level. The two thresholds:
PERC\_THRESHOLD\_LO and PERC \_THRESHOLD\_HI are empirically set.

@@

@@

\subsection{Generalizing PPF for any Prefetcher}
\label{Arch-Generalizing}
The above discussion of PPF shows that it is highly modular and can be
adapted to be used over any base prefecher for increased prefetch
accuracy.  In general, only two hooks are required between PPF and the
baseline prefetcher. The first is to make sure that all the prefetch
candidates of the prefetcher pass through the perceptron filter and if
qualified, the metadata for perceptron indexing be stored. The second
is needed when the feedback of a prior prefetch is available in form
of a subsequent demand hit or cache eviction. In that case, the stored
metadata needs to be retrieved to update the state of the perceptrons.

@@

\item \textbf{Feature Selection:} In Section~\ref{Impl-Features}, we
  show the features used for PPF. Six out of the nine features we
  developed use information derived directly from program execution,
  irrespective of the baseline prefetcher. Beyond that, the feature
  set can be expanded to convey any useful information from the
  prefetcher to the perceptron filter.  The methodology explained in
  Section~\ref{Method-Features} talks about developing a minimal
  feature set for PPF.
%For example, in our implementation ``Confidence,'' ``Current Signature XOR Delta,''
%and ``Page Address XOR Confidence'' are such features.

>>>>

As discussed in Section~\ref{sec:Background-Perceptron}, perceptrons
are useful, light-weight mechanisms to pull together disparate forms
of information and synthesize a decision from them.  In this case, we
will leverage a variety of different relative characteristics of a
given prefetch, such as speculation depth, page address and offset,
into a binary useful/non-useful prediction.  In the remainder of this
section, we discuss the the general structure of PPF for enhancing a
given underlying prefetcher.  We will show that PPF is a generalized
prefetch filtering mechanism which may adapted to any prefetcher with
minimal modifications. %Gino: Are these modifications minimal?

\subsection{The Perceptron Filter}
\label{Arch-Perceptron}
\pg{rewrite this section as talking to the figures and walking through
  them}

The Perceptron Filter is organized as tables of perceptron weights,
with a separate table of weights for each feature.  A different number
of bits are used for each feature to index into its corresponding
table.
%Gino: repharse below?
Each feature's value is used to index into the corresponding table.
This indexing is done using a different number of bits for each
feature.  At most, 12 bits of a feature are used to index into the
weights table. Certain features require more resolution \textit{i.e.},
the full 12 bits of indexing.  On the other hand, some features
requiring lower resolution require as few as 7 bits of indexing.  The
variable indexing was determined by studying the features and
fine-tuned empirically to achieve a good accuracy vs hardware overhead
trade-off.  Exact details are mentioned under ``Area Overheads'' in
Section~\ref{Method-Overheads}. \pg{this whole ``bits'' discussion
  belongs in implementation not here IMHO}

A single entry in the table corresponds to a perceptron weight.  Each
weight is a 5 bit counter - saturating at -16 and +15 and initialized
to zero at the beginning of a program's execution. Our proposed design
uses nine features, represented in nine perceptron weight
tables. \pg{the fact that its nine is again implementation details}

\pg{The word ChampSim should not appear in this part of the paper at
  all.  Remember you are describing how to build this in the general
  case...} \textbf{Inference}\newline In the ChampSim simulator, the
prefetching method is invoked on every L2 Cache demand access.  At
that point, the base prefetcher has the option to perform a prefetch
-- if it does, then it has a choice of how many cache lines to
prefetch.  The suggested prefetches can be either placed in the L2
cache or the last-level cache based on the prefetcher's internal
confidence mechanism.

@@

Once all the weights are retrieved, they are summed and compared to a
preset threshold, PERC\_ THRESHOLD\_LO.  Only the prefetch candidates
with perceptron sum higher than the threshold qualify for prefetching.
The prefetches that qualify through the perceptron stage are recorded
in the ``Prefetch Table''. The prefetch table is a 1024-entry, direct
mapped structure that contains all metadata required to recreate the
features.  %Gino: Recreate the features?
When the feedback of the current prefetched line is available at a
later stage, the stored data is used to train the perceptron.

\pg{this discussion below has redundancy and doesn't flow well.
  Rework it}

@@

The sum of the perceptron weights are also used to decide whether to
place the data in the L2 or the last level cache.  All prefetch
candidates that are qualified until this stage are compared against
PERC\_THRESHOLD\_HI to decide the fill level. The two thresholds:
PERC\_THRESHOLD\_LO and PERC \_THRESHOLD\_HI are empirically set.

@@

\subsection{Optimizing PPF for a Given Prefetcher}
\label{Arch-Generalizing}
The above discussion of PPF shows that it is highly modular and can be
adapted to be used over any base prefecher for increased prefetch
accuracy.  In general, only two hooks are required between PPF and the
baseline prefetcher. The first is to make sure that all the prefetch
candidates of the prefetcher pass through the perceptron filter and if
qualified, the metadata for perceptron indexing be stored. The second
is needed when the feedback of a prior prefetch is available in form
of a subsequent demand hit or cache eviction. In that case, the stored
metadata needs to be retrieved to update the state of the perceptrons.

@@

\item \textbf{Feature Selection:} In Section~\ref{Impl-Features}, we
  show the features used for PPF. Six out of the nine features we
  developed use information derived directly from program execution,
  irrespective of the baseline prefetcher. Beyond that, the feature
  set can be expanded to convey any useful information from the
  prefetcher to the perceptron filter.  The methodology explained in
  Section~\ref{Method-Features} talks about developing a minimal
  feature set for PPF. \pg{This should be more about how one should
    choose features than about what you did in the SPP implementation
    in particular.  Rewrite as a guide on picking features.}
%For example, in our implementation ``Confidence,'' ``Current Signature XOR Delta,''
%and ``Page Address XOR Confidence'' are such features.
